# Regression and Prediction

**Learning objectives:**

- Perform **linear regressions** with a single independent variable.
- Perform linear regressions with **multiple** independent variables.
- Perform regressions with one or more **categorical** independent variables.
- Perform **nonlinear** generalizations of regression.
- Compare and contrast the use of regression for **prediction vs. explanation**.
- **Cautiously interpret** the results of a multivariable linear regression.
- **Assess the goodness** of a regression model.

## Weighted Regression

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
```

```{r data, include=FALSE}
data(ames, package = "modeldata")
```

Used to give certain records (variables, features) more or less weighting when fitting the regression model. To show this, I will use the ames housing data from the `{modeldata}` package from `tidymodels` and prioritize sale prices of houses sold more recently than those sold earlier in these data.

```{r weighted}
dat <- ames %>% 
  dplyr::select(Lot_Area, Neighborhood, Year_Sold, First_Flr_SF, Second_Flr_SF,
         Bsmt_Full_Bath, Full_Bath, Half_Bath, Bsmt_Half_Bath, Sale_Price,
         Bedroom_AbvGr, Central_Air, Bldg_Type) %>% 
  dplyr::mutate(weight = Year_Sold - 2006,
         total_sf = First_Flr_SF + Second_Flr_SF,
         bath = Bsmt_Full_Bath + Full_Bath + 0.5*Half_Bath + 0.5*Bsmt_Half_Bath)


house_lm <- lm(Sale_Price ~ total_sf + Lot_Area + bath + Bedroom_AbvGr + Central_Air,
               data = dat)
house_wt <- lm(Sale_Price ~ total_sf + Lot_Area + bath + Bedroom_AbvGr + Central_Air,
               data = dat, weight = weight)
round(cbind(house_lm = house_lm$coefficients,
                house_wt = house_wt$coefficients), digits = 3)
```

## Prediction using Regression

**Caution: Be careful extrapolating results beyond the range of the dataset**

Prediction Interval (Uncertainty around a single value)

- Confidence Intervals (Uncertainty around a statistic)

- Individual data point error

Here is an example of individual data point error. If we filter for those properties that have four bedrooms, 3 bathrooms and a lot square footage between 10k and 11k, the sale price varies by $50k. This is error in our model.

```{r uncertainty}
dat %>% dplyr::filter(Bedroom_AbvGr == 4 & 
                 bath == 3 & 
                 Lot_Area >=10000 & 
                 Lot_Area < 11000) %>% 
  dplyr::select(Sale_Price) %>%
  dplyr::arrange(-Sale_Price)
```

## Factor Variables

We can use the building type as a factor variable to help with our predictions. The building type variable has five options: 

```{r}
dat %>% count(Bldg_Type)
```

### Dummy Variables

- One Hot Encoding (KNN, Tree Models) vs P-1 representation (Regression)

One hot encoding is when all factor levels are included in the model. *Adding in all P distinct levels along with the intercept term creates collinearity issues.*

```{r}
model.matrix(~Bldg_Type -1, data = dat) %>% head(1)
```

- P-1 encoding (Using all of the factor levels except the reference)

R uses the first factor as the reference level and we should interpret the remaining levels relative to this factor.

```{r warnings = FALSE}
lm(Sale_Price ~ total_sf + Lot_Area + bath + Bedroom_AbvGr + Central_Air + 
     Bldg_Type, data = dat) %>% 
  summary() %>% broom::tidy()
```

### Ordered Factor Variables 

Treating ordered factors as a numerical variable preserves the information contained in the ordering that would be lost if we simply used a factor conversion (Likert scales, Loan grades, Crime rate, etc).

## Interpreting Regression Equations

### Correlated Variables (Variables that move together, either in the same direction or opposite direction)

Remember back to our house sales data, the coefficient for Bedrooms was negative. This implies that the more bedrooms a house has, the less it will sell for. The reason is the total square feet and the number of bedrooms (and bathrooms) is highly correlated. We can see this using a Gaussian Graphical Model

```{r}
dat %>% dplyr::select(total_sf, Lot_Area, bath, Bedroom_AbvGr) %>% 
  correlation::correlation(partial = TRUE) %>%
  plot()
```

When we remove the total square feet and number of bathrooms, the number of bedrooms becomes desirable. We are essentially using this as a proxy for the size of the home. 

```{r}
lm(Sale_Price ~ Lot_Area + Bedroom_AbvGr + Central_Air,
               data = dat) %>% 
  summary() %>% 
  broom::tidy() 
```

### Multicollinearity (when a predictor can be expressed as a linear combination of other predictors--extreme case of correlated variables)

```{r}
lm(Sale_Price ~ total_sf + First_Flr_SF + Second_Flr_SF + Lot_Area + 
     Bedroom_AbvGr + Central_Air, data = dat) %>% 
  summary() %>% 
  broom::tidy() 
```

In the background, R handles this by removing a variable that causes multicollinearity (Second_Flr_SF). However, this is unstable and should be addressed explicitly.

### Confounding Variables (problem of ommision)

With our housing data, location may be a confounding (houses in some neighborhoods may sell at a higher price than other neighborhoods).

```{r}
neighborhood_groups <- dat %>% 
  dplyr::mutate(resid = residuals(house_lm)) %>% 
  dplyr::group_by(Neighborhood) %>% 
  dplyr::summarize(med_resid = median(resid),
            cnt = n()) %>% 
  dplyr::arrange(med_resid) %>% 
  dplyr::mutate(cum_cnt = cumsum(cnt),
        neighborhoodgroup = as_factor(ntile(cum_cnt, 5)))

dat <- dat %>% 
  left_join(dplyr::select(neighborhood_groups, Neighborhood, neighborhoodgroup),
            by = "Neighborhood")

lm(Sale_Price ~ total_sf + Lot_Area + Bedroom_AbvGr + bath + Central_Air + 
     neighborhoodgroup, data = dat) %>% 
  summary() %>% 
  broom::tidy()
```

### Main Effects and Interactions

```{r}
interaction_fit <- lm(Sale_Price ~ Lot_Area + Bedroom_AbvGr + bath + Central_Air + neighborhoodgroup*total_sf, data = dat) 

interaction_fit %>% 
  summary() %>% 
  broom::tidy()
```

If an interaction is significant, it means the association is different at different levels of a factor or different values of a continuous variable. You will need to visually determine how this differs in order to interpret these results.

```{r}
interaction <- ggeffects::ggpredict(interaction_fit, 
                                    terms = c("neighborhoodgroup", "total_sf"))

plot(interaction)
```

Selecting interaction terms

- Prior knowledge and intuition can guide choices

- Stepwise regression can be used to sift through variables

- Penalized regression can automatically fit to a large set of variables

- The most common approach is to use *tree models*, as well as their descendants which automatically search for optimal interaction terms.

## Regression Diagnostics

### Outliers (extreme value) *This may not be an influential case*

Using the `{performance}` package by Daniel Lüdecke, we identify one influential case using Cook's Distance (Other options are available).

```{r}
model <- lm(Sale_Price ~ total_sf+ bath + Lot_Area + Bedroom_AbvGr, data = dat)

outliers <- performance::check_outliers(model)
plot(outliers)

as.data.frame(outliers) %>% 
  dplyr::arrange(-Outlier_Cook) %>% 
  head()
```

It is hard to tell whether this is a typo or a one off sale. The property sold for $160k but has an almost 64k lot area and over 5.6k square footage--quite a deal in this area.

```{r}
dat %>% slice(1499) %>% select(Sale_Price, total_sf, bath, Lot_Area, Bedroom_AbvGr)
```

When we remove this influential case, our coefficients change quite a bit.

```{r}
house_noinfluence <- lm(Sale_Price ~ total_sf+ bath + Lot_Area + Bedroom_AbvGr, 
                        data = dat %>% 
                          slice(1:1498, 1500:n()))

round(cbind(house_lm = model$coefficients,
                house_noinfluence = house_noinfluence$coefficients), digits = 3)
```

### Assumption Checking (Heteroscedasticity, Normality of residuals, Linearity, and Collinearity)

```{r}
performance::check_model(model)
```

## Non-linear Regression

### Partial Residual Plots and Nonlinearity

The linearity plot gives us some indication of a non-linear fit. To dig deeper, we can look at partial residual plots using the {ggeffects} package by Daniel Lüdecke. A partial residual plot represents the residuals of one dependent and one independent variable taking into account the other independent variables.

Here is a standard scatterplot between Sales Price and Total Square Feet

```{r}
pr <- ggeffects::ggpredict(model, "total_sf [all]")
plot(pr, add.data = TRUE)
```

Here we produce a partial residual plot between Sales Price and Total Square Feet (taking into account the other independent variables). The blue line is a local polynomial regression line (loess) for reference. This indicates we may have a non-linear association.

```{r}
plot(pr, residuals = TRUE, residuals.line = TRUE)
```

### Polynomial and Spline Regression

We can create a polynomial variable (predictor squared) and add it into the model. The polynomial model seems to more accurately represent these data.

```{r}
poly_model <- lm(Sale_Price ~ poly(total_sf, 2) + bath + Lot_Area + Bedroom_AbvGr, 
                 data = dat)

polynomial <- ggeffects::ggpredict(poly_model, "total_sf")
plot(polynomial, residuals = TRUE, residuals.line = TRUE)
```

We can create a spline regression which will divides the dataset into multiple bins, called knots, and creates a separate fit for each bin. The difficult part is determining the correct knots.

```{r}
knots <- quantile(dat$total_sf, p = c(0.25, 0.5, 0.75))
lm_spline <- lm(Sale_Price ~ splines::bs(total_sf, knots = knots, degree = 3) + 
                  bath + Lot_Area + Bedroom_AbvGr, data = dat)
spline <- ggeffects::ggpredict(lm_spline, "total_sf")
plot(spline, residuals = TRUE, residuals.line = TRUE)
```

## Generalized Additive Models

```{r}
lm_gam <- mgcv::gam(Sale_Price ~ s(total_sf) + bath + Lot_Area + Bedroom_AbvGr, 
                    data = dat)
gam <- ggeffects::ggpredict(lm_gam, "total_sf")
plot(gam, residuals = TRUE, residuals.line = TRUE)
```

## Meeting Videos

### Cohort 1, Part 1

`r knitr::include_url("https://www.youtube.com/embed/URL")`

<details>
<summary> Meeting chat log </summary>

```
CHAT LOG
```
</details>
